A Runtime-Assisted Approach To Convergence in Replicated Data Types

### 2021-11-12 introduction.tex
Motivation in progress

\section{Introduction}

@@

The second distinguishing aspect of our approach is the utilization of a
distributed runtime that orchestrates \emph{well-formed} distributed
executions that are guaranteed to converge. Note that it is easy to
construct a trivial runtime that guarantees convergence of all executions.
Such a runtime would make an extensive use of inter-replica coordination to
synchronize the execution of every operation and thus induce a
sequentially-consistent behavior. This would however defeat the purpose of
state replication as the latency incurred for synchronized execution of an
operation is prohibitively high, and the availability of the system during
the execution is correspondingly low. It is therefore important for a
distributed runtime of RDTs to not synchronize the execution of RDT
operations. Orchestrating a well-formed convergent execution without
impacting latency and availability may seem impossible due to the CAP
theorem~\cite{cap}, but it is not the case in fact. While CAP theorem does
rule out linearizability, it does not prohibit enforcing weaker consistency
constraints on distributed executions. In fact many CRDTs, including the
aforementioned \emph{remove-wins} set, implicitly assume \emph{causal
consistency}, which is weaker than linearizability but is stronger than the
default \emph{eventual consistency}. While causal consistency alone cannot
guarantee convergence, it sufficiently constrains the distributed execution
for the CRDT implementations to guarantee convergence. In this respect our
  approach differs in the sense that (a). It identifies a novel set of
  constraints on distributed execution that guarantee convergence of an
  MRDT \emph{regardless} of its merge semantics, and (b). It localizes
  those constraints spatially and temporally into the runtime such that
  per-operation latency and system-wide availability remain unaffected.
Sec.~\ref{sec:semantics} explains our runtime in detail.

@@

  \item We realize an implementation of \quark as a light-weight shim layer
  on top of Scylla -- an off-the-shelf distributed key-value
  store~\cite{scylla}. An evaluation over several mergeable replicated data
  types (MRDTs) demonstrate that \quark has \emph{no} effect on the latency
  and availability of the system. To demonstrate the trade-offs not
  captured by latency and availability, we invent a new metric called
  \emph{staleness} and show that our approach does affect it negatively, albeit
  within the tolerable limits.
\end{itemize}

>>>>

\section{Introduction}
\label{sec:intro}

@@

The second distinguishing aspect of our approach is the utilization of a
distributed runtime that orchestrates \emph{well-formed} distributed
executions that are guaranteed to converge. Note that it is easy to
construct a trivial runtime that guarantees convergence of all executions.
Such a runtime would make an extensive use of inter-replica coordination to
synchronize the execution of every operation and thus induce a
sequentially-consistent behavior. This would however defeat the purpose of
state replication as the latency incurred for synchronized execution of an
operation is prohibitively high, and the availability of the system during
the execution is correspondingly low. It is therefore important for a
distributed runtime of RDTs to not synchronize the execution of RDT
operations. Orchestrating a well-formed convergent execution without
impacting latency and availability may seem impossible due to the CAP
theorem~\cite{cap}, but it is not the case in fact. While CAP theorem does
rule out linearizability, it does not prohibit enforcing weaker consistency
constraints on distributed executions. In fact many CRDTs, including the
aforementioned \emph{remove-wins} set, implicitly assume \emph{causal
consistency}, which is weaker than linearizability but is stronger than the
default \emph{eventual consistency}. While causal consistency alone cannot
guarantee convergence, it sufficiently constrains the distributed execution
for the CRDT implementations to guarantee convergence. In this respect our
approach differs in the sense that (a). It identifies a novel set of
constraints on distributed execution that guarantee the convergence of an
MRDT \emph{regardless} of its merge semantics, and (b). It localizes those
constraints spatially and temporally into the runtime such that
per-operation latency and system-wide availability remain unaffected.
Sec.~\ref{sec:semantics} explains our runtime in detail.

@@

  \item We realize an implementation of \quark as a light-weight shim layer
  on top of Scylla -- an off-the-shelf distributed key-value
  store~\cite{scylla}. We do a thorough evaluation with a
    collaborative-editing benchmark to demonstrate that \quark has
    \emph{no} effect on the latency and availability of the system, while
    slightly increasing the time to convergence.
\end{itemize}
